---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE, echo = FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(ggplot2)
theme_set(theme_bw())
```

# MEMC

[![R-CMD](https://github.com/Microbial-Explicit-Model/MEMC/actions/workflows/rcmd.yml/badge.svg)](https://github.com/Microbial-Explicit-Model/MEMC/actions/workflows/rcmd.yml)
[![codecov](https://codecov.io/gh/Microbial-Explicit-Model/MEMC/branch/main/graph/badge.svg?token=MA69VLXPYP)](https://codecov.io/gh/Microbial-Explicit-Model/MEMC)
[![Documentation](https://github.com/Microbial-Explicit-Model/MEMC/actions/workflows/pkgdown.yaml/badge.svg)](https://github.com/Microbial-Explicit-Model/MEMC/actions/workflows/pkgdown.yaml)

`MEMC` is a R package that allows users to explore various
representation of soil organic carbon (SOC) flux dynamics within a
consistent SOM model structure. Popular microbial-explicit SOC
models such as MEND^[Wang, Gangsheng, Sindhu Jagadamma, Melanie A.
Mayes, Christopher W. Schadt, J. Megan Steinweg, Lianhong Gu, and
Wilfred M. Post. 2015. “Microbial Dormancy Improves Development and
Experimental Validation of Ecosystem Model.” The ISME Journal 9 (1):
226–37.], MIMICS^[.Wieder, William R., Steven D. Allison, Eric A.
Davidson, Katerina Georgiou, Oleksandra Hararuk, Yujie He, Francesca
Hopkins, et al. 2015. “Explicitly Representing Soil Microbial Processes
in Earth System Models.” Global Biogeochemical Cycles 29 (10):
1782–1800.], and CORPSE^[Sulman, Benjamin N., Jessica A. M. Moore, Rose
Abramoff, Colin Averill, Stephanie Kivlin, Katerina Georgiou, Bhavya
Sridhar, et al. 2018. “Multiple Models and Experiments Underscore Large
Uncertainty in Soil Carbon Dynamics.” Biogeochemistry 141 (2): 109–23.]
vary significantly in their formulations of microbial mechanisms and
underlying pool structure. `MEMC` provides a consistent pool structure
with flexible flux dynamics: the conceptual and mathematical formulations
of key SOC transformations can be rapidly changed. 
This allows modelers to easily explore the
effects of different conceptualizations of microbial mechanisms.  

## Installation

Follow the download and installation instructions for 
[R](https://cloud.r-project.org/) and 
[R studio](https://www.rstudio.com/products/rstudio/download/). 

Use `remotes` to install `MEMC` as a built R package directory from GitHub. 

```
# use install.packages("remotes") to install this package the first time.
library(remotes)

# Now build and install the R package on your local machine.
install_github('Microbial-Explicit-Model/MEMC') 
```

Now load `MEMC` as you would any other package: 

```{r}
library(MEMC)
```

## Getting Started

The `MEMC` package ships with several already defined model configuration
(see `help(configurations)`). Here we demonstrate how to run a simulation
using the "MEND_config" configuration. For more examples and package
details, please check out our [online
documentation](https://microbial-explicit-model.github.io/MEMC/).
Alternatively, users can look at the `memc_all_configs` table to see all
model configurations included in the package and the microbial dynamics
used in each configuration. 

```{r}
summary(memc_all_configs)
```

Look in detail at the pre-built MEND_config configuration (see
`help("MEND_config")` for more details):

```{r}
# Printing the summary table for a MEMC model configuration, this will indicate
# the dynamics that will be used in the model run. 
summary(MEND_config)
```


Parameter value 
```{r}
# The parameter values can be view using by indexing into the MEND_config object
print(MEND_config$params) 
```


Initial pool states for when simulation time is equal to 0. 
```{r}
# Similarly the initial pool values are accessed using 
print(MEND_config$state)
```

Perform a run using the MEND model:

```{r, warning=FALSE}
time <- seq(0, 365, by = 25) 
mend_out <- memc_solve(mod = MEND_config, time = time)
```

`memc_solve` returns a long-format data frame with the state of the model
pool at each time point. This makes it easy to plot the results:

```{r}
ggplot(data = mend_out) + 
  geom_line(aes(time, value, color = name), linewidth = 0.75) + 
  facet_wrap("variable", scales = "free") + 
  labs(y = "mg C/g soil", 
       x = "time (days)",
       title = "MEND Run Results")
```

## Building a Custom Model

`MEMC` allows users to easily run simulations with the provided model
configurations and also build customized models of their own design by
selecting any combination of the supported flux dynamics, and/or modifying
the model parameters. 
For this example we will use the default parameter and initial pool
values that are included as package data (see `help("memc_params)`
and `help("memc_initial_state)` for more information).

```{r}
# Use memc_configure to print a table describing the model configuration 
my_config <- memc_configure(params = memc_params, 
                           state = memc_initial_state, 
                           name = "my model", 
                           F1 = "MM", 
                           F2 = "LM", 
                           F8 = "LM")
summary(my_config)
```

Run our customized model...

```{r, warning=FALSE}
time <- seq(0, 365, by = 25) 
my_out <- memc_solve(mod = my_config, time = time)
```

...and compare its output with the MEND model results from above: 

```{r, echo = FALSE}
ggplot(data = rbind(mend_out, my_out)) + 
  geom_line(aes(time, value, color = name), linewidth = 0.75) + 
  facet_wrap("variable", scales = "free") + 
  labs(y = "mg C/g soil", 
       title = "Comparing Model Results", 
       x = "time (days)") +
  theme(legend.title = element_blank())
```

Changing the POC and DOC decomposition flux dynamics affects model
behavior! The flexibility of the flux dynamics makes `MEMC` a powerful
tool for rapid SOM model exploration. 

Additional features supported by `MEMC`
include the ability to change model parameters, perform sensitivity
analyses, and fit models with experimental/observational data (see
[online documentation](https://microbial-explicit-model.github.io/MEMC/)
for examples featuring capabilities).
