---
title: "Example: Change Inputs"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{"Example: Change Inputs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


Here in this example we will walk through how to change the inputs to. Starting out by changing a single parameter or inital starting condition at a time. And walk through how to do a run with a custom values. This tutorial assumes that the pacakge has already been installed, [see MEMC installation instructions](installation.html). 

```{r}
library(dplyr)
library(ggplot2)
library(MEMC)
```

# `MEND`

First run the model with default values, so we that we will have something we can compare other runs with. 

```{r}
t <- seq(1,1e3)
MEND_out <- MEND(parameters = MEND_params, time = t, inital_state = MEND_initalState)
MEND_out$name <- "default MEND" # Add a column to keep track of the parameter values that were used.
```


## Change paramter values. 

Now rerun `MEND` but change the half-saturation constant value `K.d` for the Michaelis–Menten kinetics describing microbial (B) uptake of dissolved organic carbon (D). `K.d` has an inverse relationship with the uptake. When `K.d` increases the rate of D uptake by B slows, whereas when `K.d` is decreased the rate of D uptake increases. 


```{r}
# Change the values for K.d in the parameter table. 
double_kd <- mutate(MEND_params, value = if_else(parameter == "K.d", value * 2, value))
half_kd   <- mutate(MEND_params, value = if_else(parameter == "K.d", value * 1/2, value))

# Use the parameter new paramter values with `MEND`. 
MEND(parameters = double_kd, time = t, inital_state = MEND_initalState) %>% 
  mutate(name = "double K.d") ->
  MEND_out_double
  
MEND(parameters = half_kd, time = t, inital_state = MEND_initalState) %>% 
  mutate(name = "half K.d") ->
  MEND_out_half
```

Plot the results of MEND running with different half-saturation constants. 

```{r}
# Combine the data frames together
out <- rbind(MEND_out, MEND_out_double, MEND_out_half)
```

```{r}
ggplot(data = out) + 
  geom_line(aes(time, value, color = name)) + 
  facet_wrap("variable", scales = "free") + 
  theme(legend.title = element_blank()) +
  labs(title = "MEND", 
       subtitle = "varying half-saturation constant for microbial uptake",
       y = "mg C/g soil")
```


By changing the `K.d` value the rate of B uptake changes which in turn affects the rest of the system. This approach can be used to change any number of parameter values and even read in a completely new set of parameter values, so long as the parameter table meets all of `MEND` requirements. If a parameter is missing an error will be thrown. Manipulating the parameter `MEND2_params` table to change the parameter values used to solve the two pool version, `MEND2`. 


## Change inital conditions. 


```{r}
# Save a copy of default initial conditions 
state <- MEND_initalState

# Save a copy of the default value of B 
default_val <- state[which(names(state) == "B")]

# Create a copy of the inital conditions vector and double the value of B.
double_B <- state
double_B[names(double_B) == "B"] <- default_val * 2

# Create a copy of the inital conditions vector and half the value of B.
half_B <- state
half_B[names(half_B) == "B"] <- default_val * .5

# Run MEND with the different inital starting conditions. 
MEND(parameters = MEND_params, time = t, inital_state = double_B) %>% 
  mutate(name = "double B") ->
  MEND_B_double
  
MEND(parameters = MEND_params, time = t, inital_state = half_B) %>% 
  mutate(name = "half B") ->
  MEND_B_half

out <- rbind(MEND_out, MEND_B_double, MEND_B_half)

ggplot(data = out) + 
  geom_line(aes(time, value, color = name)) + 
  facet_wrap("variable", scales = "free") + 
  theme(legend.title = element_blank()) +
  labs(title = "MEND", 
       subtitle = "varying initial conditions of B",
       y = "mg C/g soil")
```


This approach can be used to change any number of initial conditions, so long as the state vector meets all of `MEND` requirements. If a value is missing an error will be thrown. Manipulating the parameter `MEND2_initalState` vector can be used to solve the two pool version, `MEND2`. 
