---
title: "MEMC"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MEMC}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, echo = FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  fig.width = 8, 
  fig.height = 5
)

library(ggplot2)
theme_set(theme_bw(base_size = 14))
```

`MEMC` is a R package that allows users to explore various
representation of soil organic matter (SOM) flux dynamics within a
consistent SOM pool model structure.

## Installation

`MEMC` may be installed using the `install_github` function from 
the `remotes` package. 

```{r}
# Install the remotes package if needed, and then
# remotes::install_github("Microbial-Explicit-Model/MEMC") 

# Load the installed MEMC package
library(MEMC)
```

## Example 1: Run `MEMC` MEND 

The package ships with several already defined model configurations; use `help(configurations)` to see a list. Currently there are six configurations 
included as internal MEMC package data, with each using different 
representations for DOM uptake, POM decomposition, and MB decay.

The MEMC model configurations are described in the `model_configs` table. 

```{r}
print(model_configs)
```

Let's try using MEND, the microbial enzyme-mediated decomposition model 
originally developed by 
[Wang et al. 2014](https://doi.org/10.1371/journal.pone.0089252).

You can type `help("MEND_model")` for more details about the pre-built configuration.

```{r}
# Printing this MEMC model configuration will return a list giving the 
# run name, a table of the flux dynamics, parameter values, and initial
# SOM pool sizes. 
print(MEND_model)
```

Perform a model run using the `memc_solve` function, whose inputs are: 

* `mod`. A `MEMC` model object, either one of the model configurations 
included with the package or created using the `memc_configure()`; and 
* `times`. A numeric vector of the time sequence for which the model will be
solved; the first value of times is the initial time. 

```{r, warning=FALSE}
mend_out <- memc_solve(mod = MEND_model, time = 0:600)
```

The model run results are saved in the `mend_out` data frame. 

```{r}
# Preview the run results
head(mend_out)
```

Visualize the run results:

```{r}
library(ggplot2)
ggplot(data = mend_out) + 
  geom_line(aes(time, value, color = name), linewidth = 0.75) + 
  facet_wrap("variable", scales = "free") + 
  labs(y = "mg C/g soil", 
       x = "time (hour)", 
       title = "MEND Run Results")
```

## Example 2: Build your own model 

With `MEMC`, users are able not only to easily run simulations with the
provided model configurations, but also build toy model of their own
design by selecting any combination of the supported flux dynamics. See
here for an example for how to use `memc_configure` to build your own
SOM model.

The `MEMC` package includes default parameter and initial pool values
that are included as package data that are based on Wang et al.
2015^[Wang, G., Jagadamma, S., Mayes, M. et al. Microbial dormancy
improves development and experimental validation of ecosystem model.
ISME J 9, 226–237 (2015). https://doi.org/10.1038/ismej.2014.120]. This
example will use these values.

Before setting up the model configuration, take a look at these defaults:

```{r}
print(memc_state_ultisol) # this is MEMC::memc_state_ultisol
```


```{r}
print(default_params) # MEMC::default_params
```

Set up the model configuration:

```{r}
# Running memc_configure will print a table describing the model configuration. 
my_model <- memc_configure(params = default_params, 
                            state = memc_state_ultisol, 
                            name = "my model", 
                            DOMuptake = "MM", 
                            POMdecomp = "LM", 
                            MBdecay = "LM")
```

Use `memc_solve` to run our model...

```{r, warning=FALSE}
my_out <- memc_solve(mod = my_model, time = 0:500)
head(my_out)
```

...and compare our toy model results with the MEND results:

```{r, echo = FALSE}
ggplot(data = rbind(mend_out, my_out)) + 
  geom_line(aes(time, value, color = name), linewidth = 0.75) + 
  facet_wrap("variable", scales = "free") + 
  labs(y = "mg C/g soil", 
       title = "Comparing Model Results", 
       x = "time (hour)") +
  theme(legend.title = element_blank())
```

For more examples and tutorials please see our [online documentation](https://microbial-explicit-model.github.io/MEMC/).
