---
title: "MEMC"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MEMC}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Installation

`MEMC` may be installed using the `install_github` function from the `remotes` package. 

```
## install the remotes package if needed
remotes::install_github("Microbial-Explicit-Model/MEMC") 
````

## Example

```{r, message=FALSE}
## load the package
library(MEMC)

## load other packages that will help with plotting results 
library(ggplot2)
theme_set(theme_bw())
```

This example uses the default package initial pool size estimated and parameter values and from Wang et al. 2015^[Wang, G., Jagadamma, S., Mayes, M. et al. Microbial dormancy improves development and experimental validation of ecosystem model. ISME J 9, 226–237 (2015). https://doi.org/10.1038/ismej.2014.120]. Before setting up the model configuration take a look at these default values. 

```{r}
print(default_initial)
# Alternatively can use the :: to access the MEMC package data. 
# print(MEMC::default_initial)
```


```{r}
print(default_params)
# Alternatively can use the :: to access the MEMC package data. 
# print(MEMC::default_params)
```

Use `configure_model` to set up the model. The `DOMdecomp`, `POMdecomp`, and `MBdecay` arguments are used to define the flux dynamics uses for DOM decomposition, POM decomposition, and Microbial Biomass decay, respectively. Where MM refers to michaelis-menten and DD refers to density dependent. 

```{r}
# Define your model using model configuration. 
my_model <- configure_model(params = default_params, 
                state = default_initial, 
                name = "MEND", 
                DOMdecomp = "MM", 
                POMdecomp = "MM", 
                MBdecay = "DD")
```

Run the model and plot results. 

```{r}
results <- solve_model(mod = my_model, # the model configuration we created above
                       time = 0:500)   # the time vector to solve the model for 
```


```{r, fig.width=8, fig.height=5}
ggplot(data = results, 
       aes(time, value)) + 
  geom_line() + 
  facet_wrap("variable", scales = "free") + 
  labs(y = "mg C/g") 
```
