---
title: "Change Model Parameters"
---

```{r, include = FALSE, echo = FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  fig.width = 8, 
  fig.height = 5
)
library(MEMC)
library(ggplot2)
theme_set(theme_bw(base_size = 14))
```

This example will show how to update the model parameter values for any MEMC model configuration. 


```{r, eval=FALSE}
# Assumes MEMC has already been installed. 
library(MEMC) 

# Use to visualize results.
library(ggplot2) 
```

The `MEMC` package includes default parameter and initial pool values that are included as package data that are based on Wang et al. 2015^[Wang, G., Jagadamma, S., Mayes, M. et al. Microbial dormancy improves development and experimental validation of ecosystem model. ISME J 9, 226–237 (2015). https://doi.org/10.1038/ismej.2014.120]. This example will use these values and show users how to update these values. We will also be running the MEMC MEND model configuration. \

# Default Parameters 

Start by taking a look at the default model parameter values. 

```{r}
print(default_initial)
# Alternatively can use the :: to access the MEMC package data. 
# print(MEMC::default_initial)
```


```{r}
print(default_params)
# Alternatively can use the :: to access the MEMC package data. 
# print(MEMC::default_params)
```

# Run deafult MEMC MEND 

Run the pre-configured MEMC MEND model with the default parameter values. 

```{r}
time <- 0:600 
mod1 <- MEND_model
out1 <- memc_solve(mod = mod1, time = time)
out1$name <- "Km = 250 (default)"
```

# Update parameter value

Users can update the parameters directly via `memc_solve`. Any parameter(s) value(s) fed into the `memc_solve` params argument will overwrite the parameter values defined the model configuration.

```{r}
out2 <- memc_solve(mod = MEND_model, time = time, params = c("K_m" = 10))
out2$name <- "Km = 10 (lower)"
```

```{r}
ggplot(data = rbind(out1, out2)) + 
  geom_line(aes(time, value, color = name, linetype = name), linewidth = 0.75) + 
  facet_wrap("variable", scales = "free") + 
  labs(y = "mg C/g soil", 
       title = "Change Km Values") +
  theme(legend.title = element_blank())
```

# Update parameter table using `update_params`

`update_params` is a helper function that can change the parameter values in the parameter table. You will need to set up a new model configuration with the new parameter table. 

```{r}
new_km <- update_params(new_params = c("K_m" = 100), param_table = default_params)

# Set up the new model configuration with the 
mod_km <- configure_model(params = new_km, 
                        state = default_initial,
                        name = "Km = 100", 
                        DOMuptake = "MM",
                        POMdecomp = "MM",
                        MBdecay = "LM")

out3 <- memc_solve(mod = mod_km, time = time)
```



```{r}
ggplot(data = rbind(out1, out2, out3)) + 
  geom_line(aes(time, value, color = name, linetype = name), linewidth = 0.75) + 
  facet_wrap("variable", scales = "free") + 
  labs(y = "mg C/g soil", 
       title = "Change Km Values") +
  theme(legend.title = element_blank())
```