---
title: "Fit model to data"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The objective of this example is demonstrate how to users can use `memc_modfit` to fit a MEMC model configuration to observational data. In this example we will be fitting the MIMCS model to the data from 
CITATION EXPERIMENT. 

Set Up 

```{r, warning=FALSE, message=FALSE}
# This assumes that the MEMC package has already been installed. 
library(MEMC)

# Used to visualize results
library(ggplot2)  

# set a theme to use in all the plots
theme_set(theme_bw() + theme(legend.title = element_blank()))
```

Start by loading the comparison data, included in the package data. 

```{r}
comp_data <- read.csv(system.file("example/Ultisol_control.csv", package = "MEMC"))
```

Set up the the MIMCS model with different initial pool sizes to be consistent with the experiment step up. 

```{r, message=FALSE, warning=FALSE}
# Save a copy of the mimics model configuration. 
my_mod_config <- MEMC::MIMCS_model

# Update the MEND model initial state conditions to reflect the experimental set up.
my_mod_config$state <- c(POM = 4.71, MOM = 17.67, QOM = 0, 
                         MB = 0.52, DOM = 0.148, EP = 0.052, 
                         EM = 0.052, IC = 0, Tot = 23.484)
```

Run the model with the original parameter values. 

```{r}
original_rslts <- memc_solve(my_mod_config, time = 0:max(comp_data$time))
```

Use `memc_modfit` to fit the model to the observed data. 

Description of inputs to MEMC function `memc_modfit` are as follows: 

* config: a MEMC model configuration object, either one of the pre-built configurations listed in model_configs or created using configure_model
* comp_data: a data frame containing the comparison data that the model will be fit to 

```{r}
fit <- memc_modfit(config = my_mod_config, 
                        comp_data = comp_data, 
                        x = c(V_d=2, V_p=5, V_m=0.01))

fit$par
```

Re run the model with the parameter values. 

```{r}
new_rslts <- memc_solve(my_mod_config, params = fit$par, time = 0:max(comp_data$time)) 
```

```{r}
# Format into a single data table to plot
original_rslts$par <- "original values"
new_rslts$par <- "fitted values"
out <- rbind(original_rslts, new_rslts)
```

Plot the model results vs the comparison data:

```{r}
ggplot(data = out[out$variable == "IC", ]) +
  geom_line(aes(time, value, color = par)) +
  geom_point(data = comp_data, aes(time, IC)) +
  facet_wrap("variable", scales = "free") + 
  labs(x = "Time (hour)", y = "mg C/g soil", title = "MIMCS") 
```
