---
title: "Install & Build"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Getting Started Using `MEMC`

Follow the download and installation instructions for [R](https://cloud.r-project.org/) and [R studio](https://www.rstudio.com/products/rstudio/download/). 

Use `remotes` to install MEMC as a built R package directory from github. 

```
# use install.packages("remotes") to install this package the first time.
library(remotes)

# Now build and install the R package on your local machine.
install_github('Microbial-Explicit-Model/MEMC') 

# You can also install from specific git tags...
install_github("Microbial-Explicit-Model/MEMC@version")

# ...branches...
install_github("Microbial-Explicit-Model/MEMC@branch_name")

# ...or commit hashes.
install_github("Microbial-Explicit-Model/MEMC@commit-hash")
```


## For developers

Clone the repository from github to install the package in development mode. Build the package by calling `devtools::load_all()` in R from the root directory. Please see [r-pkgs](https://r-pkgs.org/) for more information on how to build and develop R packages. 


After making changes to the code, checking out a new git branch, or pulling new commits you will need to do a clean rebuild of the MEMC package. From Rstudio click Build > Clean and Rebuild or run the following in the R console. 

```{r, eval=FALSE}
devtools::build()
```
