---
title: "Global Sensitivity"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The objective of this example is demonstrate how to use `memc_sensrange` to conduct a global sensitivity analysis and visualize the results to demonstrate package capabilities. 


The `memc_sensrange` function is a wrapper for `sensRange` from the FME package, designed to help `memc` users easily perform an analysis to estimate the global effect of `memc` parameter or initial condition on model output. The official [FME](https://cran.r-project.org/web/packages/FME/FME.pdf) provides more details, but in short,  `memc_sensrange` runs a memc configuration using parameter combinations drawn from some predefined distribution, running the model with each of these parameter combinations producing an envelope of model results. 


Start by loading the packages used in the demonstration. 

```{r setup}
library(MEMC) 


library(ggplot2)
theme_set(theme_bw())
```


Select the `memc` parameters or initial conditions to include in the local sensitivity analysis. Any of the parameters included in `memc_params` or `memc_initial_state` may be considered. In this example we will consider how the memc MEND configuration is sensitive to $V_d$ and $K_d$, the maximum and half saturation rates of DOM decomposition. 

Set up a data frame of the parameter range to sample from, here we arbitrarily set the upper and lower bounds as $\pm$ 50%  of the default parameter value. 

```{r}
pars <- c("V_d" = 3.0e+00,"V_p" = 1.4e+01,"V_m" = 2.5e-01)
prange <- data.frame(min = pars - pars * 0.5,
                     max = pars + pars * 1.5)
```


```{r}
MENDsens_out <- memc_sensrange(config = MEND_config,   # memc configuration
                               t = 0:100,              # the time steps 
                               x = pars,               # default parameter values
                               parRange = prange,      # parameter range to sample from
                               dist = "latin",         # sampling distribution to use, "latin" hyper cube or "norm"al 
                               n = 50)                 # the number of model runs/parameters to sample

```


The returned data frame includes the mean, standard deviation, range, and several interquartile values for the organic material pools.

```{r}
head(MENDsens_out)
```

Quickly visualize results with the ribbon bounds controlled by the lower and upper bound arguments.

```{r}
plot(MENDsens_out, lower = "Min", upper = "Max")
```

