---
title: "Senstivity-Analysis"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The objective of this example is demonstrate how to easily conduct a
sensitivity analysis and visualize results using `memc_sensrange`,
`memc_sensfunc`, and `memc_format_sensout`. This is a demonstration of
package capabilities more than a rigorous analysis. The first section
will demonstrate how to use the MEMC functions related to sensitivity
analyses, and the second will show how alternative model configurations
can affect sensitivity results.

## Getting started

```{r setup}
library(MEMC) # Load the installed MEMC package
library(ggplot2) # Used to plot results
theme_set(theme_bw())
```

## Define the paramter ranges

Set up a data frame defining the parameter ranges to sample from. Any
model parameters or initial pool sizes can be used; in this example we
explore the effects that maximum specific decomposition constants for
POM, DOM, and MOM have on model behavior. For simplicity, here we sample
a broad range (50% of the parameter value), but a real analysis would
use more informed parameter ranges from the literature,  

```{r}
pars <-  c("V_d" = 3.0e+00, "V_p" = 1.4e+01, "V_m" = 2.5e-01)
prange <- data.frame(pars = pars, 
                     min = pars * 0.50, 
                     max = pars * 1.50)
```

## `memc_sensrange`

The `memc_sensrange` function is `MEMC` specific wrapper of
`FME::sensRange`. `memc_sensrange` allows users to easily run a `MEMC`
model configuration with parameter combinations drawn from a predefined
distribution, ultimately generating an uncertainty envelope around model
results.

The inputs to MEMC's `memc_sensrange` function are: 

* `config`: a MEMC model configuration object, either one of the
pre-built configurations listed in `memc_all_configs` or created using
`memc_configure`
* `t`: vector of the time steps to run the model at 
*`pars`: vector of the parameters that will be varied 
* `parRange`: data frame of the min/max parameter values 
* `dist`: str for the distribution according to which the parameters
will be sampled from, with options "unif" (uniformly random samples),
"norm", (normally distributed random samples), "latin" (latin hypercube
distribution), and "grid" (parameters arranged on a grid)
* `n`: integer number of sampled parameter combinations to run the model
with

```{r, fig.width=6}
MENDsens_out <- memc_sensrange(config = MEND_config,
                               t = seq(0, 365), 
                               x = pars, 
                               parRange = prange, 
                               dist = "latin",
                               n = 50)
```

Take a quick look at the results. 

```{r}
head(summary(MENDsens_out))
```

Visualize results using R's base `plot` function. This is a quick way to
look at results, but for more user control over the figures we recommend
using [ggplot2](https://ggplot2-book.org). 

```{r, fig.width=8, fig.height=8}
plot(MENDsens_out)
```

```{r, fig.width=8, fig.height=8}
plot(summary(MENDsens_out))
```

## `memc_sensfun`

Given a MEMC model configuration, estimate the local effect of certain
parameters on selected sensitivity variables by calculating a matrix 
of so-called sensitivity functions based on `FME::sensFun`. 

The inputs to MEMC's `memc_sensfun` function are: 
* `config`: a MEMC model configuration object, either one of the
pre-built configurations listed in `memc_all_configs` or created using
`memc_configure`
* `t`: numeric vector of the time steps to run the model at
* `pars`: vector specifying the model parameters to test

```{r}
sens_out <- memc_sensfunc(config = MEND_config,
                          t = seq(0, 365),
                          x = c("V_d" = 3.0e+00,"V_p" = 1.4e+01,"V_m" = 2.5e-01))
```

Take a look at the results. 

```{r}
head(sens_out)
```

## `memc_format_sensout`

To visualize the results from `memc_sensfun` we use the helper function
`memc_format_sensout`, which formats the returned object into a data frame
that can be used by `ggplot`. 

```{r, message=FALSE}
ggplot(data = memc_format_sensout(sens_out)) +
  geom_line(aes(time, value, color = parameter)) +
  facet_wrap("variable", scales = "free")
```

# Global Parameter Sensivity Model Comparison 

```{r}
prange <- data.frame(pars =  c("V_d" = 3.0e+00,
                               "V_p" = 1.4e+01,
                               "V_m" = 2.5e-01), 
                     min = pars * 0.50, 
                     max = pars * 1.50)
```

Use the same parameter ranges with each of the different MEMC model configurations. 
```{r, fig.width=6}
time <- seq(0, 100)
MENDsens_out <- memc_sensrange(config = MEND_config,
                               t = time, 
                               x = pars, 
                               parRange = prange, 
                               dist = "latin",
                               n = 50)
COMISSIONsens_out <- memc_sensrange(config = COMISSION_config,
                                    t = time, 
                                    x = pars, 
                                    parRange = prange, 
                                    dist = "latin",
                                    n = 50)
CORPSEsens_out <- memc_sensrange(config = CORPSE_config,
                                 t = time, 
                                 x = pars, 
                                 parRange = prange, 
                                 dist = "latin",
                                 n = 50)
MEMSens_out <- memc_sensrange(config = MEMS_config,
                              t = time, 
                              x = pars, 
                              parRange = prange, 
                              dist = "latin",
                              n = 50)
BAMSens_out <- memc_sensrange(config = BAMS_config,
                              t = time, 
                              x = pars, 
                              parRange = prange, 
                              dist = "latin",
                              n = 50)
MIMCSens_out <- memc_sensrange(config = MIMCS_config,
                               t = time, 
                               x = pars, 
                               parRange = prange, 
                               dist = "latin",
                               n = 50)
```

Prepare all of the data into a single data frame for plotting. 

```{r}
out <- rbind(cbind(memc_format_sensout(MENDsens_out), name = "MEND"),
             cbind(memc_format_sensout(COMISSIONsens_out), name = "COMISSION"), 
             cbind(memc_format_sensout(CORPSEsens_out), name = "CORPSE"),
             cbind(memc_format_sensout(MEMSens_out), name = "MEMS"), 
             cbind(memc_format_sensout(BAMSens_out), name = "BAMS"), 
             cbind(memc_format_sensout(MIMCSens_out), name = "MIMCS"))
```

```{r}
ggplot(data = out) + 
  geom_ribbon(aes(time, ymin = q25, ymax = q75, fill = name), alpha = 0.5) + 
  geom_line(aes(time, Mean, color = name)) + 
  facet_wrap("variable", scales = "free") + 
  labs(y = "mg C/g soil", 
       x = "Time (days)",
       title = "Global Sensitivity to Vd, Vp & Vm", 
       subtitle = "q25/q75") + 
  theme(legend.title = element_blank()) + 
  scale_fill_manual(values = memc_colorPalette()) + 
  scale_color_manual(values = memc_colorPalette())
```

# Comparison of Local Parameter Sensivity 

Use the `memc_sensfun` to compare the sensitivity of the DOM pool to
`V_d` and `K_d`, two parameters that affect the MB uptake of DOM for
MEND and CORPSE. 

```{r}
time <- 0:100
MENDsens_out <- memc_sensfunc(config = MEND_config,
                              t = time,
                              x = c("V_d" = 3.0e+00,"K_d" = 0.250), 
                              sensvar = "DOM")

CORPSEsens_out <- memc_sensfunc(config = CORPSE_config,
                                t = time,
                                x = c("V_d" = 3.0e+00,"K_d" = 0.250), 
                                sensvar = "DOM")
```

Format output and visualize the results. 

```{r}
out <- rbind(cbind(memc_format_sensout(CORPSEsens_out), name = "CORPSE"), 
             cbind(memc_format_sensout(MENDsens_out), name = "MEND"))
```

```{r, fig.width=8}
ggplot(data = out) +
  geom_line(aes(time, value, color = name)) +
  facet_wrap("parameter", scales = "free") + 
  labs(title = "Local Senstivity to DOM", 
       x = "Time (days)", y = "" ) 
```

# Model Sensitivity to Inital Pool Sizes

The MEMC models are also sensitive to the initial pool size! Here we
will demonstrate how model results are sensitive to uncertainty within a
single pool! This being said the initial carbon pool sizes matter and
users will need to make sure that they are using the appropriate values
for their applications. 

```{r}
pools <-  c("DOM" = 1)
prange <- data.frame(pools = pools, 
                     min = pools * 0.50, 
                     max = pools * 1.50)
```

```{r, fig.width=6}
inital_pool_uncertainty <- memc_sensrange(config = MEND_config,
                                          t = seq(0, 365), 
                                          x = pools, 
                                          parRange = prange, 
                                          dist = "latin",
                                          n = 50)
```

```{r}
plot(summary(inital_pool_uncertainty))
```
