---
title: "Local Senstivity Analysis"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The objective of this example is demonstrate how to use `memc_sensfun` to conduct a local sensitivity analysis and visualize the results to demonstrate package capabilities. 


The `memc_sensfun` function is a wrapper for `sensFun` from the FME package, designed to help `memc` users easily perform a local sensitivity analysis on any `memc` parameter or initial condition. The official [FME](https://cran.r-project.org/web/packages/FME/FME.pdf) provides more details, but in short,  `memc_sensfun` runs a memc configuration multiple times: first as a reference run, then as a series of perturbation runs—one for each parameter of interest. During each perturbation run, a single parameter is changed slightly while others remain constant. The output of `memc_sensfun` includes the numerical approximations of partial derivatives, representing the sensitivity of the model output to each perturbed parameter.


Start by loading the pacakges used in the demonstration. 

```{r setup}
library(MEMC) 


library(ggplot2)
theme_set(theme_bw())
```

Select the `memc` parameters or initial conditions to include in the local sensitivity analysis. Any of the parameters included in `memc_params` or `memc_initial_state` may be considered. In this example we will consider how the memc MEND configuration is sensitive to $V_d$ and $K_d$, the maximum and half saturation rates of DOM decomposition. 


Create a vector of the parameter values of interest, the ones we would like to consider in the local sensitivity analysis and set equal to the the parameter values used by the `memc_config` that will be used to run `memc_sensfun`. 

```{r}
params_to_perturb <- c("V_d" =  3, "K_d" = 0.25)
```


Run the local sensitivity analysis for a memc configuration. The `config` may be any of the default configuraitons listed in `memc_all_configs` or one created by a user via `memc_configure`.

```{r}
out1 <- memc_sensfun(config = MEND_config,     
                     t = 0:100, 
                     x = params_to_perturb)
```


The returned data frame includes details about the simulation, such as its name, time, and pool. The value column contains the partial derivative of the simulation output with respect to the perturbed parameter, which is specified in the parameter column.

```{r}
head(out1)
```

Quickly plot the results to compare how parameter perturbations affect model output across all pools throughout the simulation. 

```{r}
plot(out1)
```

Now let's apply `memc_sensfun` to an other model configuration so that we may compare the results. 


```{r}
out2 <- memc_sensfun(config = CORPSE_config,     
                     t = 0:100, 
                     x = params_to_perturb)
plot(out2)
```



Figures 1 and 2 are quite different from one another, let's focus on the DOM and IC pools in a side by side comparison. But first recall what is different between these two configurations. By taking a look at their configuration tables. It appears that these models differ in how the represent DOM uptake and POM decomposition. 

```{r}
rbind(MEND_config$table, 
      CORPSE_config$table)
```


```{r}
# Save results for the two carbon pools are we are interested in. 
out1_to_compare <- out1[out1$variable %in% c("DOM", "IC"), ]
out2_to_compare <- out2[out2$variable %in% c("DOM", "IC"), ]

# Combine into a single data frame for plotting 
to_plot <- rbind(out1_to_compare, out2_to_compare)


# Plot the results 
ggplot(data = to_plot) + 
  geom_hline(yintercept = 0, alpha = 0.5) + 
  geom_line(aes(time, value, color = name, linetype = parameter)) + 
  facet_grid(variable ~ parameter, scales = "free") + 
  labs(y  = "Difference in Output", x = "Time (days)")
```
 
The figure above demonstrate how MEND and CORPSE ROM and IC pools respond differently to parameter perturbations. 





